1. Create Business Statement

2. Download tripdata for September, 2021 to August, 2022 from https://divvy-tripdata.s3.amazonaws.com/index.html

3. Call libraries that will be needed for data cleaning/importing.
```{r}
library(tidyverse)
library(lubridate)
library(skimr)
library(janitor)
library(dplyr)
```

4. Import data separated by month
```{r}
Sep21_tripdata <- read.csv("Data/202109-divvy-tripdata.csv")

Oct21_tripdata <- read.csv("Data/202110-divvy-tripdata.csv")

Nov21_tripdata <- read.csv("Data/202111-divvy-tripdata.csv")

Dec21_tripdata <- read.csv("Data/202112-divvy-tripdata.csv")

Jan22_tripdata <- read.csv("Data/202201-divvy-tripdata.csv")

Feb22_tripdata <- read.csv("Data/202202-divvy-tripdata.csv")

Mar22_tripdata <- read.csv("Data/202203-divvy-tripdata.csv")

Apr22_tripdata <- read.csv("Data/202204-divvy-tripdata.csv")

May22_tripdata <- read.csv("Data/202205-divvy-tripdata.csv")

Jun22_tripdata <- read.csv("Data/202206-divvy-tripdata.csv")

Jul22_tripdata <- read.csv("Data/202207-divvy-tripdata.csv")

Aug22_tripdata <- read.csv("Data/202208-divvy-tripdata.csv")
```

5. Check the column names to make sure they match
```{r}
colnames(Sep21_tripdata)
colnames(Oct21_tripdata)
colnames(Nov21_tripdata)
colnames(Dec21_tripdata)
colnames(Jan22_tripdata)
colnames(Feb22_tripdata)
colnames(Mar22_tripdata)
colnames(Apr22_tripdata)
colnames(May22_tripdata)
colnames(Jun22_tripdata)
colnames(Jul22_tripdata)
colnames(Aug22_tripdata)
```
6. Compare data types to avoid issues in stacking
```{r}
compare_df_cols(Sep21_tripdata,Oct21_tripdata,Nov21_tripdata,Dec21_tripdata,Jan22_tripdata,Feb22_tripdata,Mar22_tripdata,Apr22_tripdata,May22_tripdata,Jun22_tripdata,Jul22_tripdata,Aug22_tripdata, return="mismatch")
```
7. There are no mismatches, so we can consolidate the data frames into one
```{r}
tripdata_all <- bind_rows(Sep21_tripdata,Oct21_tripdata,Nov21_tripdata,Dec21_tripdata,Jan22_tripdata,Feb22_tripdata,Mar22_tripdata,Apr22_tripdata,May22_tripdata,Jun22_tripdata,Jul22_tripdata,Aug22_tripdata)
```

8. We need to remove the columns that we do not need for the analysis (Latitude and Longitude)
```{r}
tripdata_all <- tripdata_all %>% 
  select(-c(start_lat,start_lng,end_lat,end_lng))
```

9. Inspect the new dataframe
```{r}
colnames(tripdata_all)
dim(tripdata_all)
head(tripdata_all)
str(tripdata_all)
summary(tripdata_all)
```

10. Create new columns separating out the date, month, day, year, and trip duration for more granular analysis.
```{r}
tripdata_all['date'] <- as.Date(tripdata_all$started_at)
tripdata_all$month <- format(as.Date(tripdata_all$date), "%m")
tripdata_all$day <- format(as.Date(tripdata_all$date), "%d")
tripdata_all$weekday <- format(as.Date(tripdata_all$date), "%A")
tripdata_all$length_of_ride <- difftime(tripdata_all$ended_at,tripdata_all$started_at)
```

11. We need to convert length_of_ride to numeric in order to do calculations.
```{r}
tripdata_all$length_of_ride <- as.numeric(as.character(tripdata_all$length_of_ride))
```


12. Check length of ride column for inconsistencies, and make sure our previous step to convert length_of_ride to numeric worked.
```{r}
min(tripdata_all$length_of_ride)
summary(tripdata_all)
is.numeric(tripdata_all$length_of_ride)
```
13. We can see that there are values that do not make sense in the length of ride column, so we need to remove these values.  We will create a new data frame for the final cleaned data, and check to make sure we no longer have any negative numbers.
```{r}
tripdata_final <- tripdata_all[!(tripdata_all$length_of_ride < 0),]
min(tripdata_final$length_of_ride)
```

14. We need to export to a .csv file in order to use Tableau for the analysis.
```{r}
write.csv(tripdata_final, "tripdata_final.csv")
```

